---
title: "Regression Analysis"
output:   
  html_document:
    toc: true
    toc_float: true
---

```{r,include=FALSE,message=FALSE,echo=FALSE}
library(tidyverse)
library(ggplot2)
library(broom)
library(plotly)
library(highcharter)
library(rworldmap)
library(gganimate)
library(transformr)
library(corrplot)
library(leaps)
library(kableExtra)
knitr::opts_chunk$set(
 echo = FALSE,
 fig.width = 7, 
 fig.height = 5,
 fig.asp = 0.6,
 out.width = "60%")
theme_set(theme_bw() + 
          theme(legend.position = "bottom",
                legend.title = element_blank(),
                plot.title = element_text(hjust = 0.5, size = 15),
                plot.subtitle = element_text(hjust = 0.5, size = 12)))


```

### Data Import and Cleaning

#### Death dataset.

```{r}
deaths = read_csv("./data/final_hiv_deaths.csv") %>% 
  janitor::clean_names() %>% 
  select("location_name", starts_with("sex"), starts_with("age"), "year", "val", "upper", "lower") %>% 
  rename(country_name = location_name)

deaths %>% 
  summarize(n_countries = n_distinct(country_name),
            n_years = n_distinct(year),
            col_name = colnames(deaths),
            n_missing = map(deaths, ~sum(is.na(.)))) %>% 
  unnest(n_missing)
```

The death dataset comes from:
It contains HIV death count data from 204 distinct countries, in the 30 years between 1990 and 2019.

#### GDP and population datasets.

```{r}
gdp = read_csv("./data/final_gdp_per_capita.csv", skip = 4, col_names = TRUE) %>% 
  pivot_longer(cols = "1960":"2021",
               names_to = "year",
               values_to = "GDP per capita") 

population = read_csv("./data/final_population_total.csv", skip = 4, col_names = TRUE) %>% 
  pivot_longer(cols = "1960":"2021",
               names_to = "year",
               values_to = "Population") 

gdp_population = inner_join(gdp, population, by = c("Country Name", "year")) %>% 
  janitor::clean_names() %>% 
  select("country_name", "year", "gdp_per_capita", "population") %>% 
  mutate(year = as.numeric(year))

gdp_population %>% 
  summarize(n_countries = n_distinct(country_name),
            n_years = n_distinct(year),
            col_name = colnames(gdp_population),
            n_missing = map(gdp_population, ~sum(is.na(.)))) %>% 
  unnest(n_missing)
```

The GDP and population datasets both come from:
After combining, the gdp_population dataset contains data from 266 distinct countries, in the 62 years between 1960 and 2021.

#### Merging datasets to produce the "maindata" dataset for analysis purpose.

```{r}
maindata = left_join(deaths, gdp_population, by = c("year","country_name"))

maindata %>% 
  summarize(n_countries = n_distinct(country_name),
            n_years = n_distinct(year),
            col_name = colnames(maindata),
            n_missing = map(maindata, ~sum(is.na(.)))) %>% 
  unnest(n_missing)
```

By matching on the countries and years from the "death" dataset using "left_join", we produced the main dataset. It has a size of `r nrow(maindata)` observations and `r ncol(maindata)` features. Basic features we are interested in include:

- **country_name**: 
- **sex_id**
- **sex_name**
- **age_id**
- **age_name**: Age groups including "0-9 years", "10-24 years",  "25-49 years", "50-74 years", and "75+ years".
- **year**: Year of the records
- **val**: 
- **upper**: 
- **lower**: 
- **gdp_per_capita**:

```{r}

```

population by year
```{r}
pop_plot = gdp_population %>% 
ggplot(aes(x = population))  
```

#### Description

### Global trend of HIV mortality per 100k population overtime (1990 to 2019)

```{r}
death_year = maindata %>% 
  group_by(year) %>% 
  summarize(HIV_death = sum(val),
            pop_year  = sum(population, na.rm = T)) %>% 
  mutate(
    HIV_death = as.integer(HIV_death)
  ) %>% 
  mutate(
    HIV_death = as.numeric(HIV_death),
    motality_per_100k = HIV_death/(pop_year/100000)) 

death_year %>% 
  ggplot(aes(x = year, y = motality_per_100k, color = year)) + geom_point()
```


### Global trend of HIV mortality per 100k population by gender overtime (1990 to 2019)

```{r}
death_gender = maindata %>% 
  group_by(year, sex_name) %>% 
  summarize(HIV_death = sum(val),
            pop_year  = sum(population, na.rm = T)) %>% 
  mutate(
    HIV_death = as.integer(HIV_death)
  ) %>% 
  mutate(
    HIV_death = as.numeric(HIV_death),
    motality_per_100k = HIV_death/(pop_year/100000))

death_gender %>% 
  ggplot(aes(x = year, y = motality_per_100k, color = sex_name)) + geom_point()
```

### Global trend of HIV mortality per 100k population by age overtime (1990 to 2019)

```{r}
death_age = maindata %>% 
  group_by(year, age_name) %>% 
  summarize(HIV_death = sum(val),
            pop_year  = sum(population, na.rm = T)) %>% 
  mutate(
    HIV_death = as.integer(HIV_death)
  ) %>% 
  mutate(
    HIV_death = as.numeric(HIV_death),
         motality_per_100k =     HIV_death/(pop_year/100000))
death_age %>% 
  ggplot(aes(x = year, y = motality_per_100k, color = age_name)) + geom_point()
```

### HIV mortality by Country across years

### Top 10 countries with highest HIV mortality

Add a map in this chunk (top 10 countries change overtime)? More current time? 
```{r}
death_country = maindata %>% 
  group_by(country_name) %>% 
  summarize(HIV_death = sum(val, na.rm = T),
            pop_year  = sum(population, na.rm = T)) %>% 
  mutate(
    HIV_death = as.integer(HIV_death)
  ) %>% 
  mutate(
    HIV_death = as.numeric(HIV_death),
         motality_per_100k =     HIV_death/(pop_year/100000))

top_10_high = death_country %>% 
  filter(motality_per_100k != 'Inf') %>% 
  arrange(desc(motality_per_100k)) %>% head(n = 10) %>% select(country_name,motality_per_100k)

knitr::kable(top_10_high)

```

### Top 10 countries with lowest HIV mortality

```{r}
top_10_low = death_country %>% 
  filter(motality_per_100k != 'Inf') %>% 
  arrange(motality_per_100k) %>% head(n = 10) %>% select(country_name, motality_per_100k)

knitr::kable(top_10_low)
```

### Statistical analysis

```{r}
regression = maindata %>% 
  mutate(motality_per_100k = val / (population/100000),
         log_motality_per_100k = log(motality_per_100k))

regression %>% 
  summarize(col_name = colnames(regression),
            n_missing = map(regression, ~sum(is.na(.)))) %>% 
  unnest(n_missing)

```
### Plots of the distribution to main outcome of interest(HIV mortality per 100k population).
```{r}
regression %>% 
  ggplot(aes(x = motality_per_100k)) +
  geom_histogram(color = "darkblue", fill = "lightblue") +
  ggtitle("Histogram of HIV deaths per 100k") +
  geom_vline( aes(xintercept = mean(motality_per_100k)),
             linetype = "dashed") +
  ylab("") +
  xlab("Estimated HIV deaths per 100k")


regression %>% 
  ggplot(aes(x = log_motality_per_100k)) +
  geom_histogram(color = "darkblue", fill = "lightblue") +
  ggtitle("Histogram of log(HIV deaths per 100k)") +
  geom_vline(aes(xintercept = mean(log_motality_per_100k)),
             linetype = "dashed") +
  ylab("") +
  xlab("Estimated log(HIV deaths per 100k)")
```

### 
```{r}
```


```{r}
```


```{r}
```

